\chapter*{Preface} % Main chapter title
\addcontentsline{toc}{chapter}{Preface}
\label{preface}

Physics is a wonderful thing, I was always mesmerized by its ability to describe what surround us in a powerful and compact way, ultimately using this knowledge for the betterment of mankind. When I was young, equations were like an ancient magic language to me. I was not able to understand it yet, but I was absolutely impressed by knowing that people that could, were able to control the power that for a kid look like beyond any form of comprehension, from making incredibly heavy object flying faster than any bird or destroying an entire city by splitting something invisible for the naked eye. I always dreamed at some point to become part of this exclusive club of magicians, and now that I am, I have to say: I got it all wrong! Physics requires a lot of thinking, a lot of reading, a lot of trial and error, in one word? A lot of work! What follows is my attempt to make a lasting contributes to this amazing field. It won't go to history, but I hope it will be important for some students that follow the scientist step, for sure it will be important to me.

So what is this thesis about? Dark Matter, if I had to use one word. One of the most prominent puzzles in all physics. With all our powerful and extremely precise models, we still have to explain more than 96\% of the matter in our universe. That is embarrassing! How could we miss all that? Turns out is not simple at all when the matter that you are searching for stubbornly refuses to interact with your detectors, but I am sure that eventually, physics will prove to be even more stubborn in their measurement, and crack the code. This thesis was an attempt to this, an experiment to produce this elusive matter directly, and then measure its property. Since my group does not have a noble price in its hand, it must be no mystery that we did not yet succeed, although this is not excluded for the future. I think the journey of me and the rest of my colleagues is nevertheless very instructive, it is the typical story of how an experiment is born and conducted. A "happy ending" is not always expected, and should not be assumed by the scientist that is conducting the experiment, you are just there to observe, not decide. In my opinion, the experiment itself is the "happy ending" that a scientist is seeking, when it is well performed, and that is not depending on the outcome of it. So with no further wait, we can begin our journey, at the discovery of the NA64 experiment and its search of Dark Matter using the SPS\footnote{Super Proton Synchrotron}.

In this first chapter, our story begins, and as every story of science, it begins with a mystery: Dark Matter. Why can't we see it? Why can't we touch it? Those are not interesting questions, since many well-understood phenomena can't be seen or touched. In the end, it all boils down to one question: Why our models, that works so amazingly well in so many different situations, fail miserably in other situations apparently equivalent? This is what this chapter will be about, understand why we need the concept of dark matter, and justify what phenomena it could explain. Indeed one could think that is rather lazy to explain phenomena just by adding invisible matter in the system. "Just admit that you don't know what is going on!", is something that I heard myself when I try to explain my work to others. That is overall an honest question, a healthy scientific skepticism about a theory that seems so arbitrary. In Sec.\ref{ch1:sec:dm-evidence} we will explore why we need this concept, and what are the alternatives to it. After that, we will come to realize that Dark Matter is a very simplistic term, that hides an extremely large number of possibilities that can be constructed using the framework of Quantum Field Theory (QFT). Dark Matter candidates are indeed very numerous, but we can focus on the most reasonable ones, that do not require very complicated models. In agreement with the scientific method, we try to extend our theory minimally, to see if it works, and only after we fail we add more complexity to our new model. In Sec.\ref{ch1:sec:dm-candidates} we will give a small review on these possibilities, trying to be succinct but complete. After this introduction, we will present a more specific model, which is one of the main focus of this thesis: the $U'(1)$ model. This model postulates the existence of an additional $U(1)$ symmetry that generates a "Dark Sector" of particles amounting to the invisible matter. The vector gauge boson generated by this symmetry, that we will label $\DM$ through this thesis, has been known with the catchy name of Dark Photon, due to playing the equivalent role of the standard model photon in this new symmetry. This model allows also a cross-term that couples the new Dark Photon with its standard counterpart, effectively building a portal between the two sectors. This is excellent news! It means we can produce this type of matter using modern accelerators if such a model is true, which allows us to perform a particle physics experiment to probe this model. But why this model should be better than the others? There is not an easy answer to this. From a scientific point of view, a model is better than another if it is more powerful in explaining the reality of things. In a framework as complicate as cosmology however, countless models are equivalent in describing reality, which is why so many Dark Matter candidates exist. In a way, it is unavoidable that we share the work and probe all models one by one. However, sometimes a model can solve more than one problem at the same time, which makes it especially attractive for scientists. Originally, the Dark Photon was also an excellent explanation of the measured value of the anomalous muon magnetic moment, that to date deviates substantially (by $\sim$3.6 standard deviations) from the prediction of QED. We now know that such an explanation is ruled out since NA64 together with other experiments excluded it. Another interesting phenomenon that could be explained in this framework is the so-called $\DMX$-anomaly, originally known with the name of $^8$Be-anomaly. This name refers to a bump detected in the nuclear decay spectrum of Beryllium that would be justified by the existence of a particle not present in the standard model. In Sec.\ref{ch1:sec:dm-u1model-motivations-x17} a review of this phenomena will be given. To avoid confusion, we point out here that the $\DMX$, which is the name given to the particle advocated to explain the observed anomaly, is not by definition a Dark Photon, since it posses some properties that are not present in the original model. However, its characteristics are similar enough to be probed by the NA64 experiment, as it will be clear from this thesis.

Chapter \ref{chapter1} is meant to just answer the question: what are we searching for? This is the beginning of every scientific experiment, but arguably the easiest part! Building an experiment to produce the Dark Photon is the next step, which will be covered in chapter \ref{chapter2}. Turn a photon in a Dark Photon using this portal is however meaningless if we can't prove that we did it! A robust analysis method of the data collected needs to be performed for this purpose, this is what we will explore in chapter \ref{chapter3} step by step, from the method to the selection criteria. In chapter $\ref{chapter4}$ we will then provide what most physicists are here for, the results! No Dark Matter has been found yet, but the data acquired allows us to exclude some specific models from being a credible explanation of reality. To reiterate an example already used: we now know thanks to this data, that the anomalous magnetic moment of the muon cannot be explained exclusively bu the $U'(1)$. The NA64 experiment is however far from over! The stop caused by the LHC long shutdown (and unfortunately by the recent rise of Covid-19 as well) of the accelerator allowed us to look for ways to improve our setup for the upcoming challenges of 2021! The pieces of knowledge we gained for the background and the experimental condition are used to design the new version of the setup to acquire a larger number of particles and thus probe a larger number of models. In chapter \ref{chapter5} we will review all these changes, in particular the new setup for the visible mode of 2021, that was the last project of my Ph.D., and that will use to hopefully find (or else exclude) the $\DMX$.

With no further wait, let's embark on this journey!


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../PhDthesis"
%%% End:
